---
title: "Processing DTMs for Harvard Forest"
author: "Nate Looker"
date: "June 23, 2016"
output: html_document
---

## Process Harvard forest extent kml
```{r kml-to-shp, engine='sh', eval=F}
ogr2ogr -f 'ESRI Shapefile' /home/look/Documents/data/NEONDI-2016/NEONdata/D01/harv_boundary.shp /media/look/AOP-NEON1-4/D01/HARV/2014/HARV_L1/HARV_Lidar/BoundaryKMLs/2014_HARV_2_v02_full_boundary.kml 
```

## Load libraries

```{r load-lib}
library(raster)
library(rgdal)
library(maptools)
library(cowplot)
library(data.table)
library(dplyr)
set.seed(0)
```


## Read in site boundary shapefile

```{r reproj-shp}
dir_HARV_loc <- "../NEONdata/D01/" # local data directory
dir_HARV_sd <- "/media/look/AOP-NEON1-4/D01/HARV/2014/" # data directory on NEON hard drive
harv_bound <- shapefile(paste(dir_HARV_loc, "harv_boundary.shp", sep=""))

newproj <- projection(raster("/media/look/AOP-NEON1-4/D01/HARV/2014/HARV_L3/HARV_Lidar/DTM/2014_HARV_2_720000_4712000_DTM.tif"))

harv_UTM <- spTransform(harv_bound, CRS(newproj))

plot(harv_UTM)
```

## Read in national hydrography dataset shapefiles

```{r read-in-NHD}
# Read in HUC12
harv_nhd <- shapefile(paste(dir_HARV_loc, "NHD/Shape/WBDHU12.shp", sep=""))
plot(harv_nhd)
harv_nhd_utm <- spTransform(harv_nhd, CRS(newproj))

harv_nhd_clip <- crop(harv_nhd_utm, extent(harv_UTM))
plot(harv_nhd_clip)

# Read in waterbodies
harv_wb <- shapefile(paste(dir_HARV_loc, "NHD/Shape/NHDWaterbody.shp", sep=""))

harv_wb_utm <- spTransform(harv_wb, CRS(newproj))
harv_wb_clip <- crop(harv_wb_utm, extent(harv_UTM))
plot(harv_wb_clip)
shapefile(harv_wb_clip, filename="../outputs/HARV/harv_wb_clip.shp", overwrite=T)

```

```{r read-dem-tif}

harv_dtm <- raster(paste(dir_HARV_loc, "HARV-provisional/2014/lidar/HARV_lidarDTM.tif", sep=""))


```

```{r pheno-site-extent, eval=F}

pheno_bound <- data.frame(ID=1:2, t(matrix(c(42.541779, -72.173509, 42.543716, -72.170825), ncol=2)))
names(pheno_bound) <- c("ID","Y", "X")
pheno_bound <- pheno_bound[,c(1,3,2)]
coordinates(pheno_bound) <- c("X", "Y")
projection(pheno_bound) <- projection(harv_nhd)
pheno_bound_utm <- spTransform(pheno_bound, CRS(newproj))
plot(pheno_bound_utm)
# extent(pheno_bound_utm)

```


## Clip DTM by phenology plot extent

```{r clip-dtm, eval=F}

harv_dtm_sub <- crop(harv_dtm, extent(pheno_bound_utm))
plot(harv_dtm_sub)
writeRaster(harv_dtm_sub, "../outputs/HARV/harv_dtm_sub.tif", overwrite=T)
#writeRaster(harv_dtm_sub, "../outputs/HARV/harv_dtm_pheno.sdat")

```

## Read in saved version of cropped dtm so that knitr doesn't break

```{r write-dtm-crop}

harv_dtm_sub <- raster("../outputs/HARV/harv_dtm_sub.tif")

```

## Call SAGA through bash

```{r call-saga-cmd, engine='sh', eval=F}

saga_cmd ta_morphometry 0 -ELEVATION ~/Documents/data/NEONDI-2016/outputs/HARV/harv_dtm_pheno.sgrd -SLOPE ~/Documents/data/NEONDI-2016/outputs/HARV/harv_slp -ASPECT ~/Documents/data/NEONDI-2016/outputs/HARV/harv_asp



```



## Bring in dtm from SJER

```{r read-sjer-dtm}

dir_SJER_loc <- "/home/look/Documents/data/NEONDI-2016/NEONdata/D17-California/SJER/"

sjer_dtm <- raster(paste(dir_SJER_loc, "2013/lidar/SJER_lidarDTM.tif", sep=""))
plot(sjer_dtm)

```


## Assess effects of aggregating 1-m DTMs
```{r aggregate-dtms}
## Create hillshade raster directly from DEM
hillshade_direct <- function(dem) {
  slp <- terrain(dem, opt="slope", unit="radians", neighbors=8)
  asp <- terrain(dem, opt="aspect", unit="radians", neighbors=8)
  shade <- hillShade(slope=slp, aspect=asp, angle=45, direction=315)
  return(shade)
}

## Compare hillshades from aggregated and non-aggregated DEMs
harv_sub_mean <- aggregate(harv_dtm_sub, fact=3, fun=mean)

harv_sub_mean_shade <- hillshade_direct(harv_sub_mean)
harv_sub_1m_shade <- hillshade_direct(harv_dtm_sub)

par(mfrow=c(1,2))
plot(harv_sub_1m_shade)
plot(harv_sub_mean_shade)

```

## Reclassify rasters

```{r rast-reclass}

  ## Function to reclassify a raster using a vector of quantiles (default is quartiles)
  bin_rast <- function(raster_in, probs=seq(0,1,length=5)) {
  raster_in[raster_in==0] <- NA
  breaks <- unname(quantile(raster_in, probs))
  raster_in <- cut(raster_in, breaks=breaks)
  return(raster_in)
}

```


## Read in SAGA topographic wetness tifs

```{r saga-twi}
sjer_twi <- raster("../SAGA_NEON/saga_twi_sjer.tif")
harv_twi <- raster("../SAGA_NEON/saga_twi_harv.tif")

sjer_twi_df <- values(sjer_twi)
sjer_twi_df <- sjer_twi_df[complete.cases(sjer_twi_df)]
harv_twi_df <- values(harv_twi)
harv_twi_df <- harv_twi_df[complete.cases(harv_twi_df)]

twi_df <- rbindlist(list(data.frame(Site="SJER",sjer_twi_df), data.frame(Site="HARV", harv_twi_df)))
names(twi_df)[2] <- "TWI"
rm(sjer_twi_df);rm(harv_twi_df)

density(sjer_twi, main="Topographic Wetness Index\n SJER")
density(harv_twi, main="Topographic Wetness Index\n HARV")
```

## Read in veg rasters

```{r chm-and-ndvi}

sjer_chm <- raster("../NEONdata/D17-California/SJER/2013/lidar/SJER_lidarCHM.tif")
sjer_evi <- raster("../NEONdata/D17-California/SJER/2013/spectrometer/veg_index/SJER_EVI.tif")
sjer_ndvi <- raster("../NEONdata/D17-California/SJER/2013/spectrometer/veg_index/SJER_NDVI.tif")
sjer_pri <- raster("../NEONdata/D17-California/SJER/2013/spectrometer/veg_index/SJER_PRI.tif")

harv_chm <- raster("../NEONdata/D01/HARV-provisional/2014/lidar/HARV_lidarCHM.tif")

```

## Create "Huber index" (a proposed proxy for Huber value)

```{r huber}

sjer_huber <- sjer_evi / sjer_chm
cellStats(sjer_huber, median)
sjer_huber[sjer_huber > 0.2 | sjer_huber < 0] <- NA
plot(sjer_huber)
writeRaster(sjer_huber, "../outputs/SJER/sjer_huber.tif")

```

## Reclassify TWI

```{r twi-recl}

sjer_twi_recl <- bin_rast(raster_in = sjer_twi, probs=seq(0,1,length=10))
sjer_twi_recl <- crop(sjer_twi_recl, extent(sjer_huber))
plot(sjer_twi_recl)
# writeRaster(sjer_twi_recl, "../outputs/SJER/sjer_twi_recl.tif", overwrite=T)

harv_twi_recl <- bin_rast(raster_in = harv_twi, probs=seq(0,1,length=10))
harv_twi_recl <- crop(harv_twi_recl, extent(harv_chm))

```


## Associate veg data with TWI classes

```{r ext-hub-twi}
sjer_brick <- brick(sjer_evi, sjer_chm, sjer_twi_recl, sjer_ndvi, sjer_pri)
sjer_brick_df <- as.data.frame(sjer_brick)
sjer_brick_df <- sjer_brick_df[complete.cases(sjer_brick_df),]
sjer_brick_df <- sjer_brick_df[sample(1:nrow(sjer_brick_df), 10000), ]

harv_brick <- brick(harv_chm, harv_twi_recl)
harv_brick_df <- as.data.frame(harv_brick)
harv_brick_df <- harv_brick_df[complete.cases(harv_brick_df),]
harv_brick_df <- harv_brick_df[sample(1:nrow(harv_brick_df), 10000), ]

```

## Summarize veg data by TWI class

```{r aggregate-veg}
s.e. <- function(vec) {
  sd(vec)/length(vec)
}
sjer_summary <- sjer_brick_df %>% mutate(TWI=layer) %>% group_by(TWI) %>% summarise(EVI = mean(SJER_EVI), CHM = mean(SJER_lidarCHM), seEVI = s.e.(SJER_EVI), seCHM = s.e.(SJER_lidarCHM), NDVI = mean(SJER_NDVI), seNDVI = s.e.(SJER_NDVI), PRI = mean(SJER_PRI), sePRI = s.e.(SJER_PRI))
harv_summary <- harv_brick_df %>% mutate(TWI=layer) %>% group_by(TWI) %>% summarise(CHM = mean(HARV_lidarCHM), seCHM = s.e.(HARV_lidarCHM))
twosite_summary <- rbind(data.frame(Site="SJER", sjer_summary[,c(1,3,5)]), data.frame(Site="HARV", harv_summary))
plot_out <- ggplot(twosite_summary, aes(TWI, CHM, group=Site)) + geom_point(aes(size=seCHM)) + geom_smooth(se=F, colour=1) + facet_wrap(~Site, scales="free_y") + labs(x="Topographic Wetness Index (binned by deciles)", y = "Canopy Height Model (m)")

ggsave("../outputs/figs/fig_ntl.png", plot_out)

plot_out_2 <- ggplot(sjer_summary, aes(TWI, EVI)) + geom_point(aes(size=seEVI)) + geom_smooth(se=F, colour=1)  + labs(x="Topographic Wetness Index (binned by deciles)", y = "Enhanced Vegetation Index")

ggsave("../outputs/figs/fig2_ntl.png", plot_out_2)

ggplot(sjer_summary, aes(TWI, NDVI)) + geom_point(aes(size=seNDVI)) + geom_smooth(se=F, colour=1)  + labs(x="Topographic Wetness Index (binned by deciles)", y = "Normalized Difference Vegetation Index")

ggplot(sjer_summary, aes(TWI, PRI)) + geom_point(aes(size=seNDVI)) + geom_smooth(se=F, colour=1)  + labs(x="Topographic Wetness Index (binned by deciles)", y = "Photochemical Reflectance Index")

```




